# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE} 
raw = as.data.frame ( read.csv (unz("activity.zip", "activity.csv")) )

activity = subset(raw,!is.na(raw$steps))

head(activity)
unlink("activity.csv")
```

## What is mean total number of steps taken per day?

```{r echo=TRUE}
mean(activity$steps)
median(activity$steps)
hist(activity$steps)
```

## What is the average daily activity pattern?

```{r echo=TRUE}
ag = aggregate(steps ~ interval, data=activity, FUN=mean)
plot (ag, type="l")

index_of_max = which.max(ag$steps)

```

The ```r ag$interval[index_of_max]```-th 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps - ```r ag$steps[index_of_max]```.

## Imputing missing values
```{r echo=TRUE}
missing = subset(raw,is.na(raw$steps))
n = nrow(missing)
```

The number of missing values of steps is ```r n```.

Now replacing the missing values with the mean values of particular time interval.

```{r echo=TRUE}
newActivity = raw
for (i in 1:nrow(raw)) {
  row = raw[i,]
  if (is.na(row$steps)) {
    interval = row$interval
    idx = which (ag$interval==interval)
    newActivity$steps[i] = ag$steps[idx]    
  }
}
head(newActivity)
mean(newActivity$steps)
median(newActivity$steps)
hist(newActivity$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Note: The following analysis is based on the data with imputed missing values.

```{r echo=TRUE, fig.width=6, fig.height=8}

w = cbind(newActivity, weekdays(strptime(newActivity$date, "%Y-%m-%d")))
colnames(w)[4] <- "Day"
weekdays = subset (w, !(Day %in% list("Sunday", "Saturday")))
weekends = subset (w, Day %in% list("Sunday", "Saturday"))

agWeekdays = aggregate(steps ~ interval, data=weekdays, FUN=mean)
agWeekends = aggregate(steps ~ interval, data=weekends, FUN=mean)
par(mfrow=c(2,1)) # 2 plots in 1 column   
par(xpd=TRUE)
plot (agWeekdays$steps, type="l", main="Weekdays", ylab="Avg. number of steps", xlab="5-minute interval")
plot (agWeekends$steps, type="l", main="Weekends", ylab="Avg. number of steps", xlab="5-minute interval")

```
